\textls[-0]{\thetitle}

### 2023-03-04 metatheory-condensed.tex
minor

\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT
(and \cref{fig:fmltt-selected} \TT).
%We briefly review the base MLTT fragment first.
%
Following the current trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are in the form of \citeauthor{debruijn1964} indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Substitutions are typed with the form 
%\EDJ{It seems to me G |- g : D is a judgement not a rule. Rules usually have two parts (premises + conclusion). Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here). I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this. In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
%
The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
where $\SubstId$ is the identity substitution (\ruleref{sub/id}).

>>>>

\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT
(and \cref{fig:fmltt-selected} \TT).
%We briefly review the base MLTT fragment first.
%
In keeping with the recent trend~\cite{altkap2016,coquand2018canonicity,gratzer-sterling-birkedal-2019}
in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are in the form of \citeauthor{debruijn1964} indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Substitutions are typed with the form 
%\EDJ{It seems to me G |- g : D is a judgement not a rule. Rules usually have two parts (premises + conclusion). Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here). I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this. In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
%
The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
where $\SubstId$ is the identity substitution (\ruleref{sub/id}).

### 2023-03-04 related-work.tex
make all comments into pdf comment from reviewer.
resolve most comments
minor

\citet{delaware2011} engineer product lines of theorems and proofs built
from feature modules. Feature composition is manual, which seems to have
motivated later approaches based on \emph{data types à la carte} (DTC)~\cite{dtc}. % Reviewer B : Line 910
DTC has in turn served as a source of inspiration for numerous work
\cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
%\cite{vdr2022intrinsically}
The original DTC encoding requires type-level general recursion that
fails the strict positivity check imposed by proof assistants including Coq and Agda.
\citet{delaware2013} introduce \emph{meta-theory à la carte} (MTC):
it overcomes the problem by using Church encodings for
data types and using Mendler-style folds for evaluation,
though it requires \lsti{Set} impredicativity.
Feature composition is automated through heavy use of type classes.
The framework is implemented as a Coq library.
\citet{schwaab2013modular} adapt DTC to Agda by considering a restricted
class of functors that admit least fixed points.
\citet{keuchel2013generic} use datatype-generic programming techniques
for the underlying representation of type-level fixed points and
avoid \lsti{Set} impredicativity.

>>>>

\citet{delaware2011} engineer product lines of theorems and proofs built
from feature modules. Feature composition is manual, which seems to have
motivated later approaches based on \emph{data types à la carte} (DTC)~\cite{dtc}.
%\rB{ define abbreviation DTC at first use?}

%DTC is as a source of inspiration for may approaches
%\cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
The original DTC encoding requires type-level general recursion that
fails the strict positivity check imposed by proof assistants including Coq and Agda.
\citet{delaware2013} introduce \emph{meta-theory à la carte} (MTC):
it overcomes the problem by using Church encodings for
data types and using Mendler-style folds for evaluation,
though it requires \lsti{Set} impredicativity.
Feature composition is automated through heavy use of type classes.
The framework is implemented as a Coq library.
\citet{schwaab2013modular} adapt DTC to Agda by considering a restricted
class of functors that admit least fixed points.
\citet{keuchel2013generic} use datatype-generic programming techniques
for the underlying representation of type-level fixed points and
avoid \lsti{Set} impredicativity.\YZ{What can we say about Intrinsically-Typed Definitional Interpreters a La Carte?}%\cite{vdr2022intrinsically}
